---
title: "Assigment_5"
output:
  pdf_document: default
  html_document: default
  word_document: default
date: "2024-11-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 5 by Raj Kapadia, Dhruv Daiya, Bansari Mehta, Dhruvi Patel

Loading the required libraries and packages

```{r}
library(ISLR)

source("D:\\Uni Resources\\Data Science\\Stats Inference\\A5\\regboot1.pck")
source('D:\\Uni Resources\\Data Science\\Stats Inference\\A5\\regday2.pck')
source('D:\\Uni Resources\\Data Science\\Stats Inference\\A5\\fixregday1.pck')

weather <- NOAAGISSWD
default <- Default

```

Running the logitboot function on the Wildfires

```{r}
logitboot(Wildfire.Count~delta.temp, weather, 5000)
```

No zeroes present in this

```{r}
 logitboot(Wildfire.Count~Year, weather, 5000)

```

No zeroes present in this

```{r}
 logitboot(Wildfire.Count~I(Year + delta.temp), weather, 5000)

```

No zeroes present in this

```{r}
 logitboot(Wildfire.Count~I(Year * delta.temp), weather, 5000)
```

No zeroes present in this

```{r}
 logitboot(Wildfire.Count~I(Year^2), weather, 5000)

```

No zeroes present in this

```{r}
 logitboot(Wildfire.Count~I(delta.temp^2), weather, 5000)

```

No zeroes present in this

```{r}
logitboot(Wildfire.Count~delta.temp + Year + I(delta.temp * Year) + I(delta.temp^2) + I(Year^2), weather, 5000)
```

Since there were a lot of warnings while running the above function, there is a possibility of having zeroes in them

Now running the same logitboot function but on Drought.Count

```{r}
 logitboot(Drought.Count~delta.temp, weather, 5000)

```

There might be zeroes in this function

```{r}
 logitboot(Drought.Count~Year, weather, 5000)

```

There might be zeroes when using this function

```{r}
 logitboot(Drought.Count~Year + delta.temp, weather, 5000)

```

There are zeroes when this function is used

```{r}
 logitboot(Drought.Count~I(Year * delta.temp), weather, 5000)

```

This function has no zeroes when implemented

```{r}
 logitboot(Drought.Count~I(Year^2), weather, 5000)

```

No zeroes when this function is used

```{r}
 logitboot(Drought.Count~I(delta.temp^2), weather, 5000)

```

Since there were error messages when this function was used, this indicates the possible presence of zeroes when this function is used

```{r}
 logitboot(Drought.Count~delta.temp + Year + I(delta.temp * Year) + I(delta.temp^2) + I(Year^2), weather, 5000)
```

Since there were lot of warning messages, there is a possibility of there being zeroes upon using this function

For Drought.Count only the graphs of Year \* Delta and Year\^2 had no zeroes present in their intercepts, and out of those two Year \^2 had the lowest AIC.

**Now computing the Severe Storm Count by implementing lmboot function on them**

```{r}
 lmboot(Severe.Storm.Count~delta.temp, weather, 5000)

```

This is a usable model since there are no zeroes present when using this function, which gives a PRESS of 308.5 on 95% Confidence interval

```{r}
 lmboot(Severe.Storm.Count~Year, weather, 5000)
```

This is also a usable model since there are no zeroes present when using this function, which gives a PRESS of 278.5 on 95% Confidence interval.

```{r}
 lmboot(Severe.Storm.Count~I(Year + delta.temp), weather, 5000)

```

This is also a usable model since there are no zeroes present when using this function, which gives a PRESS of 277.4 on 95% Confidence interval.

```{r}
 lmboot(Severe.Storm.Count~I(Year * delta.temp), weather, 5000)

```

This is also a usable model since there are no zeroes present when using this function, which gives a PRESS of 304.3 on 95% Confidence interval.

```{r}
 lmboot(Severe.Storm.Count~I(Year^2), weather, 5000)

```

This is also a usable model since there are no zeroes present when using this function, which gives a PRESS of 276.8 on 95% Confidence interval.

```{r}
 lmboot(Severe.Storm.Count~I(delta.temp^2), weather, 5000)

```

This is also a usable model since there are no zeroes present when using this function, which gives a PRESS of 277.5 on 95% Confidence interval.

```{r}
 lmboot(Severe.Storm.Count~Year + delta.temp + I(Year^2) + I(delta.temp^2), weather, 5000)

```

This is not a usable function, since upon running it against Severe Storm Count, there are zeroes present which gives the PRESS of 154.8 on 95% Confidence interval.

```{r}
lmboot(Severe.Storm.Count~Year + delta.temp + I(delta.temp*Year) + I(Year^2) + I(delta.temp^2), weather, 5000)
```

This is not a usable function, since upon running it against Severe Storm Count, there are zeroes present which gives the PRESS of 143.6 on 95% Confidence interval.

Running the lmboot model on All Disasters

```{r}
 lmboot(All.Disasters.Count ~ delta.temp, weather, 5000)

```

This is not a usable model, since when ran against All Disasters Count, there is zero present, which gives us a PRESS of 651.2 on 95% Confidence interval.

```{r}
 lmboot(All.Disasters.Count ~ Year, weather, 5000)

```

This is a usable model, since when upon implementing this function, there were no zeroes present, which gave us a PRESS of 597.2 on 95% Confidence interval.

```{r}
 lmboot(All.Disasters.Count ~ I(Year + delta.temp), weather, 5000)

```

This is a usable model, since when upon implementing this function, there were no zeroes present, which gave us a PRESS of 594.8 on 95% Confidence interval.

```{r}
 lmboot(All.Disasters.Count ~ I(Year * delta.temp), weather, 5000)

```

This is not a usable model, since when ran against All Disasters Count, there is zero present, which gives us a PRESS of 643.3 on 95% Confidence interval.

```{r}
 lmboot(All.Disasters.Count ~ I(Year^2), weather, 5000)

```

This is a usable model, since when upon implementing this function, there were no zeroes present, which gave us a PRESS of 594.2 on 95% Confidence interval.

```{r}
 lmboot(All.Disasters.Count ~ I(delta.temp^2), weather, 5000)

```

This is a usable model, since when upon implementing this function, there were no zeroes present, which gave us a PRESS of 492.6 on 95% Confidence interval.

```{r}
 lmboot(All.Disasters.Count ~ Year + delta.temp + I(delta.temp*Year) + I(Year^2) + I(delta.temp^2), weather, 5000)
```

This is not a usable model, since when ran against All Disasters Count, there is zero present, which gives us a PRESS of 333.7 on 95% Confidence interval.

There were many models that were usable for comparing against Severe Storm and All Disaster. There were some functions that even after removing 1 from them, they still didn't work.

The lowest PRESS for Severe Storm 143.6 when used (Year + delta.temp + I(delta.temp\*Year) + I(Year\^2) + I(delta.temp\^2), weather, 5000)

The lowesr PRESS for All Disaster Count was 492.6 which was obtained from (Temperature \^2)
