---
title: "Assignment7"
output:
  html_document: default
  pdf_document: default
date: "2024-11-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Team Member: Dhruv Daiya, Raj Kapadia, Dhruvi Patel, Bansari Mehta

```{r}
library(tidyverse)

crop_data <- read.csv("D:\\Uni Resources\\Data Science\\Stats Inference\\A7\\crop.data.csv")

# Convert `block`, `density`, and `fertilizer` to factors as they are categorical variables
crop_data$block <- as.factor(crop_data$block)
crop_data$density <- as.factor(crop_data$density)
crop_data$fertilizer <- as.factor(crop_data$fertilizer)

# Perform a multiway ANOVA including main effects and interactions
anova_model <- aov(yield ~ block + density + fertilizer + density:fertilizer, data = crop_data)


anova_summary <- summary(anova_model)
print("ANOVA Summary:")
print(anova_summary)
```

### ANOVA Results:

-   **Main Effects**:

    -   **Block**: Significant at p = 0.0016, indicating that the different blocks have a significant impact on yield

    -   **Fertilizer**: Highly significant at p = 0.0003, indicating that the type of fertilizer has a significant effect on yield.

    -   **Density: Fertilizer Interaction**: Not significant (p = 0.535), meaning there is no evidence to suggest that the interaction between density and fertilizer affects the yield.

    -   **Residuals**: These represent the variation in the data not explained by the factors in the model:

        -   Sum of Squares = 29.851.

        -   Mean Square = 0.339.

```{r}
tukey_results <- TukeyHSD(anova_model)

print("Tukey's HSD Results:")
print(tukey_results)
```

### Tukey's HSD Results:

**Significant Pairwise Comparisons**:

-   **Block**:

    -   **Block 2 - Block 1:** p = 0.037, confidence interval [ 0.020 , 0.901 ]. Block 2 has a significantly higher yield than Block 1.

    -   **Block 4 - Block 3:** p = 0.035, confidence interval [ 0.023 , 0.904 ]. Block 4 has a significantly higher yield than Block 3.

    -   **Non-significant differences:**

        -   **Block 3 - Block 1**: p=0.828p = 0.828p=0.828

        -   **Block 4 - Block 1**: p=0.235p = 0.235p=0.235

        -   **Block 4 - Block 2**: p=0.836p = 0.836p=0.836

-   **Fertilizer**:

    -   **Fertilizer 3 - Fertilizer 1:** p = 0.0003, confidence interval [ 0.252 , 0.946 ]. Fertilizer 3 significantly increases yield compared to Fertilizer 1.

    -   **Fertilizer 3 - Fertilizer 2:** p = 0.013, confidence interval [ 0.076 , 0.770 ]. Fertilizer 3 significantly increases yield compared to Fertilizer 2.

    -   **Fertilizer 2 - Fertilizer 1:** p = 0.451, non-significant.

Fertilizer 3 shows significant differences compared to the others.

-   **Density: Fertilizer Interaction**:

    -   **1:3 - 1:1:** p = 0.013, confidence interval [0.096,1.296]. The interaction of density 1 with fertilizer 3 increases yield significantly compared to density 1 with fertilizer 1.

    -   **2:3 - 1:1:** p = 0.018, confidence interval [ 0.075 , 1.275 ]. The interaction of density 2 with fertilizer 3 increases yield significantly compared to density 1 with fertilizer 1.

    -   **Non-significant differences for other interaction pairs:**

        -   **2:1−1:1**,p=0.9592:1 - 1:1, p = 0.9592:1−1:1,p=0.959

        -   **1:2−1:1**, p=0.5721:2 - 1:1, p = 0.5721:2−1:1,p=0.572

        -   **2:2−1:1**, p=0.9442:2 - 1:1, p = 0.9442:2−1:1,p=0.944

        -   **1:3−1:2**, p=0.5121:3 - 1:2, p = 0.5121:3−1:2,p=0.512

Only two interactions show significant differences.

## 

## Conclusion:

All the three main effects are not significant, block and fertilizer are significant but Density Fertilizer interaction is not significant

No significant interactions were detected in the ANOVA model. Tukey’s HSD reveals a few significant pairwise differences in the interaction terms, but they do not represent an overall significant interaction effect.

**Pairs of means that show significance under Tukey's HSD:**

-   **Block:** Differences between Block 2 and Block 1, and Block 4 and Block 3 are significant.

-   **Fertilizer:** Fertilizer 3 shows significant differences compared to Fertilizers 1 and 2.

-   **Density: Fertilizer Interaction:** Only two specific pairwise differences ( 1 : 3 − 1 : 1 1:3−1:1 and 2 : 3 − 1 : 1 2:3−1:1) are significant.
